Discovery of Rotational Modulations in the Planetary-mass Companion 2M1207b: A Slow Rotation Period and Cloud Height in a Low Gravity Atmosphere

### 2015-09-15 summary/letter/letter_vAD.tex
its been a long time since last commitment

\title{Variability of Planetary Mass Companion 2M1207 B}
\shorttitle{Variability of 2M1207 B}
\author{Yifan Zhou, Daniel Apai, Glenn Schneider}
\affil{Steward Observatory, the University of Arizona}
\author{Mark Marley}
\affil{NASA Ames Research Center}

\begin{abstract} Rotational modulations in disk-integrated light of
brown dwarfs have recently provided powerful constraints on the
properties of ultracool atmospheres, including longitudinal and
vertical cloud structures and cloud evolution. Furthermore, detection
of periodic light curve variations can directly probe the rotational
periods of ultracool objects.

We present here, for the first time, time-resolved high-precision
photometric measurements of a planetary-mass companion, 2MASS1207b, to
a brown dwarf primary. Using HST/WFC3 and point spread function %in
combination with two spacecraft roll angles we detect photometric
modulations in the light curve. The amplitude is 0.9\% in the F160W
and 1.5\% in the F125W filters; we find a consistent period and
similar phase in both bands. Joint fit to the lightcurve in both bands
suggest a period of $10.2^{+0.9}_{-0.8}$ h. The relative amplitudes in the two
filters are very similar to that found in a recent study of a field
(high-gravity) L-dwarf, suggesting that the cloud structures that
introduce the photometric modulations are similar in high- and
low-gravity objects. Importantly, our study also measures, for the
first time, the rotational period for directly imaged planetary-mass
companion.
\end{abstract}

@@

Cloud properties in high- and low-gravity objects -- thick clouds: an
explanation for very red and faint near-infrared fluxes of directly
imaged planets

Rotational mapping is powerful: successes in brown dwarfs...

The target 2M1207b... contrast The key challenge of obtaining the
lightcurve of 2M1207b is the high contrast to and small separation
from 2M1207A.

In this Letter we present the first high-contrast, high-precision,
time-resolved observations of a directly imaged planet or
planetary-mass object. We successfully detect rotational modulation
and measure the amplitudes in two bands and determine the rotational
period.

\section{Observation} We obtained direct images of the 2M1207A+b
system on UT 2014 April 11from 08:07:47 to 16:53:18 using the Hubble
Space Telescope (HST) and its Wide Field Camera 3 \citep[WFC3,
][]{Kimble2008} in the frame of the HST Proposal GO-13418 (PI:
D. Apai). We acquired the observations in filters F125W
($\lambda_{\mbox{pivot}}$ = 1245.9 nm, full width at half maximum
(FWHM) = 301.5 nm) and F160W ($\lambda_{\mbox{pivot}}$ 1540.52, FWHM =
287.9 nm), roughly corresponding to the J and H bands. The WFC3 pixel
scale is $\approx$13mas. We used the $256\times256$ pixels sub-array
mode to avoid memory dumps during the observations.  In order to
provide a near-continuous coverage for detecting modulations we
observed the 2M1207 system in six consecutive HST orbits, obtaining
data with cadence of $\sim1.5$ minutes over a baseline of 8 hours and
40 minutes. The observations were interrupted by 58 minute-long Earth
occultations every 94 minutes.

The observations applied space craft rolls between each two orbits to
allow roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbits 1, 3, and 5, and those taken in
orbits 2, 4, and 6 differed by $25^{\circ}$. At the separation of
2M1207b this angle difference corresponds to a displacement of
$0.34''$ or 2.75 and 2.30 resolution elements in F125W and F160W,
respectively. In each orbit we took thirteen SPARS10 non-destructive
read-outs with NSAMP=10, alternating between F160W and F125W filters,
with 2--3 identical exposures in one exposure sequence. To improve
sampling and reduce the risk that point spread funciton (PSF) lays on
bad pixels, we applied a 4-point dither patternwith differential "X/Y"
offsets of 1.375" in the detector frame, providing optimal
non-integral (half pixel) stepping of 10.5 and 8.5 pixels in F125W and
F160W, respectively.. Over the 6 orbits, we obtained 70 images with 10
non-destructive read-outs in F125W and 64 images in F160W with
exposure time of 88.4~s in each filters.

@@

%The small angular separation of 2M1207 A and B (as shown in Figure
%\ref{fig:1}) makes precise primary star subtraction and photometry
%very difficult. 
%On the under-sampled WFC3 IR detector (plate scale
  %$\sim 0.13''\mbox{pixel}^{-1}$ \cite{dressel2012wide}), the primary
%and the secondary only separate by $\sim 6$ pixels, which is about 5
%times of the FWHM of the PSF. In addition, under-sampling causes
%significant artifacts when shifting the images for registration no
%matter what interpolation method is used.

@@

For both of 2 rounds, we used {\em Tiny Tim} to calculate 10$\times$
over-sampled model PSFs based on the filters, the spectra 
\citep{Bonnefoy2014, Patience2010}, the telescope's actual focus, and
the telescope jitter.  We used the new set of Tiny Tim parameters
provided by \cite{Biretta2014} to improve model of the cold mask,
OTA spikes, and the coma. The focus parameters were calculated
using the model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}. To
register \tinytim{} PSF to observed PSF of 2M1207A, we moved the
over-sampled PSF over a coordinate grid (gird size=0.001 pixel) using
cubic interpolation and minimize the rms difference of the
observed and re-binned \tinytim{} PSFs over a region centered on 2M1207A with a
5-pixel-radius aperture centered on 2M1207b excluded. Then we introduce
another \tinytim{} PSF for 2M1207b and  fit the
position of 2M1207 b and the photometries of 2M1207 A and b simultaneously
by least square optimization. From our data set, we discovered that
the difference of observed PSFs and model PSFs were very stable for
given PSF positions. Therefore at the end of the first round PSF
fitting, we derived 8 (2 roll angles $\times$ 4 dithering positions)
observationally derived correction maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{obs.} - PSF_{model} )}
\end{equation}
where $\mathrm{PSF_{model}}$ is a combination of two \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combined the correction
term linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and fitted the three components together. We found that by including the correction term,
the reduced $\chi^{2}$ of PSF fitting is decreased from $\sim 10$ to
around unity. Relative photometry is acquired from the scaling
parameters of the model PSFs.

PSF profiles change with exposure positions due to
pixelation, especially for the case that WFC3 IR is significantly
under-sampled. Also, the flat fields may potentially have large scale
structures \citep{dressel2012wide}. We found a correlation of
photometry with PSF positions on detector frame for both 2M1207 A and
b. Correction were made by normalizing each group of
exposures that have the same dithering position and roll angle
individually -- we took the median of the fluxes that were measured
from these exposures as normalization factors and divided them from
every photometric measurement. Because the normalization factor for
each group of exposures is calculated across the whole observation,
this normalization step have negligible impacts on variability
analysis.

% {\em Hybrid PSF Calculation:} The hybrid PSFs were generated as a
% linear combination of a diffraction-limited PSF (calculated with the
% {\em Tiny Tim} simulator\citep{Krist1995}, and a empirically
% determined correction map. We used {\em Tiny Tim} to calculate model
% PSFs for the filters used, the spectrum of our target, the telescope's
% actual focus, and the telescope jitter.  We used the new set of Tiny
% Tim parameters provided by \cite{Biretta2014} to better model the cold
% mask, diffraction spikes, and the coma. The focus parameters were
% calculated using the model listed on the STScI
% website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}.
% A significant advantage of Tiny Tim PSF over empirical PSFs is that
% Tiny Tim can produce Nyquist-sampled results, which allows accurate
% image shifting and interpolation. However, an important limitation of
% Tiny Tim is that the model PSFs have significant systematic
% errors. \cite{Biretta2014} demonstrated that the diffraction spikes
% and coma are not accurately simulated with Tiny Tim for WFC3/IR
% images. However, using the 6-orbit long time series, we are able to
% accurately characterize the difference of the modeled and observed
% PSFs and derive an empirical correction map: Corr = Mean (
% PSF\_modeled - PSF\_observed ) where PSF\_modeled is a combination of
% two Tiny Tim PSFs with positions and amplitudes optimized to minimize
% the residuals.  By combining the model PSFs with the empirical
% correction map we derived a high-quality hybrid PSF, the use of which
% significantly improved the PSF subtraction and photometry.

% {\em PSF Fitting:} In the next step we simultaneously fitted the positions and amplitudes of the combination of the hybrid PSFs to 2M1207A and 2M1207b. First, we determined the primary star position and telescope jitter. We produced a list of 10$\times$ over-sampled PSFs matching the spectrum \citep{Bonnefoy2014} and the position of 2M1207A with different telescope jitters ranging from 0 to 50$\mbox{mini-arcsec s}^{-1}$ along
% both $x$ and $y$ detector axes. The PSFs were registered with the original image by searching on a dynamic grid and minimizing the difference in the observed and modeled PSFs at a region centered on 2M1207 A with a 5-pixel-radius aperture centered 2M1207 B excluded.

% Next, we determined the position and approximate flux of 2M1207B by calculating and fitting a Tiny Tim PSF for the hybrid PSF-subtracted 2M1207 B based on its position on the detector, spectrum \citep{Patience2010}, and
% the telescope jitter that had been obtained above. In the final step,
% we combined the two PSFs together.

% We fixed the position of the primary as it had been fitted in the first step and fitted for the
% amplitudes of the two PSFs and the precise position of the
% secondary. Since the total fluxes of the model PSFs were normalized to
% unity as default, the fluxes of 2M1207 A and B were solely represented
% by the amplitude of the two PSFs coordinately.

% Because of the systematic errors of the Tiny Tim PSF, the quality of the
% fit is not perfect. The reduced $\chi^{2}$ values were at level
% of $\sim10$. However, we found that the residual patterns were very
% stable among the image that have the same dithering position and
% telescope rolling. Therefore, we median combined the residual images
% that have the same dithering position and telescope rolling angle to
% construct 8 residual models (4 dithering position $\times$ 2 telescope
% rolling angle) for each filter. We pre-subtract the corresponding
% residual from the original image and repeated the procedure that
% listed above. Using this extra-step, the reduced $\chi^{2}$
% greatly decreased to around unity.

 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
  variability verification tests. Individual measurements are plotted with gray
  crosses. Photometries of the same exposure sequence are binned, and
  binned photometry are  plotted with points or squares. Best fitted sinusoidal
  curves are plotted with solid lines. {\em Upper}: binned photometries taken
in dithering position 1 and 3 (red points) and that taken in 2 and 4
(blue squares) are plotted with different symbols. They demonstrate same trend of
variation. In upper left panel, green line is sinusoidal wave fitted with all
parameter set free, and purple line is sinusoidal wave fitted with period set the same as that
of F125W. {\em
  Middle}: sinusoidal waves fitted without using the data taken in
Orbit \#1. These curves are almost identical to the curves
plotted in upper panel. {\em Lower}: photometry measured with
AFEM-added images and best fitted sinusoidal curves. These photometries and curves are also almost identical to
those plotted in the upper panel.}
  \label{fig:2}
\end{figure*}

\subsection{Uncertainty Analysis: White noise}
First we estimated the photon noises for the photometry of 2M1207b. We
propagated the photon noises of every single pixel that were
calculated from count rates and detector gain, to the PSF fitting
results. The photon noises for photometry in F125W and F160W are
1.33\% and 1.02\%, respectively.

Since the PSFs for the 2M1207 A and b were fitted simultaneously in our
PSF photometry procedure, the uncertainties of photometry and position
for the primary and secondary are coupled. Imperfection of position
measurement of 2M1207A could potentially affect the photometry of
2M1207b. We used a Monte Carlo (MC) method to
evaluate the overall systematic of the PSF fitting. We applied the
PSF photometry to images that were added with random Poisson noises and
repeat the photometry procedure for 1000 times. From the distribution of the
result, the uncertainty for F125W and F160W photometry are 
1.34\% and 1.12\%, respectively. We conclude that the white noise of
our observation is dominated by photon noise.

@@

2M1207b were observed at 8 different spots on the
detector (2 rolls $\times$ 4 dithering positions). Imperfect flat
field correction are potential sources of variation as they may
introduce position-dependent differences in the count rates. The
uncertainty of WFC3 IR pipeline flat field is typically $\sim 1\%$
\citep{dressel2012wide}.
% and our time resolved observations for another target
%demonstrate that aperture photometry for exposures taken at different
%dithering position are uncertain at the same level.
%Flat field uncertainties have smaller effects on 
In PSF photometry, however, multiple pixels are fitted simultaneously
and we expect a lower than 1\% uncertainty from the flat field
errors. To verify this we multiplied every image by an artificial flat
field error mask (AFEM) -- a uniformly distributed Gaussian noise array with
mean of 1 and sigma of 1\% -- and repeated the PSF photometry on the
resulting images.  The analysis of these experiments showed almost
identical lightcurve to the original, verifying that the flat field
errors do not affect our photometry significantly (Figure
\ref{fig:2}).

@@

 As a second test we repeated the analysis neglecting the first
 orbit. The motivation behind this test was that, due to spacecraft
 thermal settling, the first orbit of HST observations is often
 slightly unstable and is often neglected in high-precision studies
 (\textbf{citation?}). Indeed, in our analysis 2M1207 A is
 significantly fainter in the first orbit (Figure \ref{fig:3}) than in
 the subsequent ones.
 %The exposure levels at the pixels of the peak of the PSF of 2M1207 B are less than 10,000 e$^{-}$
 %for both two colors, thus image persistence is less of an issue for
 %2M1207 B. 
 Our analysis based on orbits 2--6 only found essentially identical
 results to our analysis of orbits 1--6, based on which we conclude
 that the first less reliable orbit does not affect our results
 significantly (Figure \ref{fig:2}, middle panel).

@@

 % \begin{itemize}
% \item the source of the scattering -- 
% \item flat field uncertainty -- adding artificial flat field error mask
% \item validity of the variability\\
%   -- both filter demonstrate similar period of variability\\
%   -- split the data into two half, the trend in two halves of data
%   looks similar.
%   -- fix the position does not change the light curve
% \item uncertainty of sinusoidal curve fit
%   -- carry out a mcmc fit? is it worth doing this.
% \end{itemize}

@@

\label{Results}
We present the first high-contrast, high-resolution, high-cadence, and
high-precision photometry of a directly imaged planet or
planetary-mass companion. Our observations reveal a modulation in the
light curve of the 5--7~M$_{J}$ companion 2M1207b, the first detection
of modulations in directly imaged ultracool objects.  The best fit
periods for F125W and F160W are 10.5 and 9.1 hour correspondingly. The
amplitudes for the normalized light curves are 1.45\% and 0.92\% for
F125W and F160W light curves.

@@

% In the following we will discuss the amplitude and period of 2M1207B
% and place this object in the broader context of ultracool atmospheres.

The distributions for the periods demonstrate long tail shaped towards
long period, with core region roughly Gaussian. With 64\% confidence,
we estimated the 1-$\sigma$ range for the periods of F125W and F160W
to be $10.5_{-1.2}^{+1.3}$ and $9.1_{-1.0}^{+1.1}$ h,
respectively. The period of best fitted sine wave of F125W light
curve is 1.5h longer than that of F160W  that is slight larger
1-$\sigma$ standard deviation. We also jointly fit the two band light
curve together forcing the periods of two sine waves to be the
same. We derived a modulation period of $10.2^{+0.9}_{-0.8}$ h in this
circumstance.

We discovered that  the variation amplitudes in the two bands were significantly
different. The distributions of the amplitudes are well described by
Gaussian profiles. By fitting a Gaussian function to the distribution, we
determined that amplitude distribution of F125W peaks at 1.45\% with a
standard deviation of 0.22\%, and that of F160W have mean and standard
deviation of 0.92\% and 0.20\%, respectively. The peaks of the two
histograms separated by more than 2-$\sigma$. The variation amplitude
of F125W light curve is 1.58 times of that of F160W light curve.

%Although the photometric time series demonstrate relatively large
%scattering, both F125W and F160W light curves demonstrate sinusoidal
%shape with clear temporal variability. We fitted sine waves to the two
%light curves using least square optimization, and determined the
%periods and variation amplitudes.

@@

\begin{figure*}
  \centering
  \plottwo{rotationDiagram}{JH}
  \caption{comparison of 2M1207's rotation period and color change with brown
    dwarfs, \bpic{} b, and solar system planets. {\em Left}: period vs. mass
  plot for 2M1207b (red square), solar system planets and \bpic{} b
  (blue squares), and brown dwarfs (black circles, gray shade). The mass of brown
  dwarfs are assumed to be 30 \mjup{}. The gray rectangle
  ranging from 15 \mjup{} to 50 \mjup{} in $x$, and $\pm\sigma$
  with the center at mean rotation period of brown dwarfs in $y$, indicates a
region where brown dwarfs most likely to appear in this diagram. {\em Right}: ratio
of variation amplitude in J and H band vs. spectral type for 2M1207b
and brown dwarfs. Points are
color coded with their J$-$H colors. }
  \label{fig:5}
\end{figure*}

A fundamental result of our study is the direct determination of the
rotation period of a directly imaged planetary-mass object. In
Figure~\ref{fig:5} we compare the rotation period of 2M1207b to the
solar system planets, field brown dwarfs that have well defined
rotation period from the study of
\citep[][]{Metchev2015}, and \bpic{} b the only other directly imaged
planet with an estimated period from measured \vsini{}.  The study by
\citep[][]{Snellen2014} succeeded in mesuring \vsini{} for \bpic{} b
and demonstrated that it fits a trend defined by Solar System planets
in which more massive planets have faster rotation rates. The
interesting finding that \bpic{} b, an exoplanet that formed in a
protoplanetary disk, follows this trend suggests a possibly connection
between planet mass, initial angular momentum, and formation in a
disk.

Excitingly, our measurement of the rotation period of 2M1207b, an
object with similar mass and age to \bpic{} b, show that their rotation
periods are very similar, {\em in spite } of the fact that 2M1207b 
most likely not formed in a circumstellar (or circumsubstellar) disk
but is a result of gravitational fragmentation. This suggests
that rotation periods are not good tracers of the formation pathways
and may not contribute important evidence for a formation in a disk
vs. in a cloud core environment.

Our observations, furthermore, allow us to compare the relative
amplitudes in the J- and H-bands between the handful of brown dwarfs
for which high-quality near-infrared time-resolved observations have
been obtained. In the right panel of Figure \ref{fig:5}, we plot the
relative amplitude of J- and H-bands of 2M1207b with brown dwarfs
\citep{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2014} that
have different spectral types and J$-$H colors.

We found a strong correlation between the spectral type of the object
and the J to H variation amplitude ratio. Figure \ref{fig:5}
demonstrates that earlier spectral type objects -- independent of
their surface gravity -- have larger variations at shorter wavelength
than at longer wavelengths. Importantly, although the J$-$H color of
2M1207b is significantly redder the other L5 dwarfs, we note that its
relative variation amplitude ratio is almost identical to the matching
spectral type mid-L dwarf 2M1821. This suggests that the upper cloud
layers, probed through variability, may be more influenced by spectral
type and less by cloud thickness, which is thought to influence the
J$-$H color.

>>>>

\title{Discovery of Rotational Modulations in the Planetary-mass
  Companion 2M1207b: A Slow Rotation Period and Cloud Height in a Low
  Gravity Atmosphere}
\shorttitle{Variability of 2M1207 B}
\author{Yifan Zhou, Daniel Apai, Glenn Schneider}
\affil{Steward Observatory, the University of Arizona}
\author{Mark Marley}
\affil{NASA Ames Research Center}

\begin{abstract}
  Rotational modulations of brown dwarfs have recently provided
  powerful constraints on the properties of ultra-cool atmospheres,
  including longitudinal and vertical cloud structures and cloud
  evolution. Furthermore, detection of periodic light curve variations
  can directly probe the rotational periods of ultra-cool objects.

  We present here, for the first time, time-resolved high-precision
  photometric measurements of a planetary-mass companion, 2MASS1207b,
  to a brown dwarf primary. Using HST/WFC3 and point spread
  function %in
  combination with two spacecraft roll angles we detect photometric
  modulations in the light curve. The amplitude is 0.9\% in the F160W
  and 1.5\% in the F125W filters; we find a consistent period and
  similar phase in both bands. Joint fit to the light curve in both
  bands suggest a period of $10.2^{+0.9}_{-0.8}$ h. The relative
  amplitudes in the two filters are very similar to that found in a
  recent study of a field (high-gravity) L-dwarf, suggesting that the
  cloud structures that introduce the photometric modulations are
  similar in high- and low-gravity objects. Importantly, our study
  also measures, for the first time, the rotational period for
  directly imaged planetary-mass companion.
\end{abstract}

@@

% Cloud properties in high- and low-gravity objects -- thick clouds: an
% explanation for very red and faint near-infrared fluxes of directly
% imaged planets
Presence of condensate clouds is one the most unique and essential
features of the ultra-cool atmosphere of direct imaged exoplanet and brown
dwarfs. The formation and properties of condensate clouds of are
intensively studied in recent years \citep[e.g.][]{Ackerman2001,
  Helling2008, Allard2012}, and great improvement has been
achieved in understanding the cloud behavior across different spectral
types, especially in the explanation of L-T transition \citep{Marley2010}.
Surface gravity is suggested to be another key parameters 
in defining cloud structures. Low surface gravity objects (e.g. HR8799 bcd, 2M1207
b) have significantly redder color compared to their spectral type
matched brown dwarfs, which are best explained by the existence of 
thick clouds \citep{Barman2011, Skemer2011, Skemer2012}. However, due to lack of observational constraint, the
behavior of clouds in the atmospheres of low surface gravity objects
are not as well as modeled as high surface gravity counterpart.

Intensity modulations introduced by heterogeneous clouds can be
directly observed and studied via time resolved observation and
rotational mapping. These techniques isolate the effect of cloud
properties and obtained great success in determining the rotation
period and unveiling the structures of the atmosphere of brown dwarfs
\citep[e.g.][]{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2014,Metchev2015}. \cite{Kostov2013}
demonstrated the great potential of these techniques in the study of the cloud properties of
direct imaged exoplanets and planetary mass companions. However, high
brightness contrast makes it much more challenging for direct
imaged exoplanet and planetary mass companions than brown
dwarfs to acquire high-precision light curves.

2M1207 b \citep{Chauvin2005} is a planetary mass companion that has mass of
$\sim 4M_{\mathrm{Jup}}$ and an angular separation of $0.78''$ to its host
2M1207 A,
an M8 brown dwarf. Relative moderate contrast (6.7 mag in J band, \cite{Mohanty2007}) makes it one
the most feasible planetary mass companions for high-precision
time-resolved observation and the best starting points for developing
strategies for time-resolved observation and data reduction in high
contrast case.

In this {\em Letter} we present the first, high-cadence, high-precision,
time-resolved {\em Hubble Space Telescope} (HST) photometric time
series of 2M1207 b, a directly imaged planet or
planetary-mass object. We successfully detect rotational modulation
and measure the amplitudes in two bands and determine the rotational
period.

\section{Observation}

We obtained direct images of the 2M1207A+b
system on UT 2014 April 11from 08:07:47 to 16:53:18 using HST and its Wide Field Camera 3 \citep[WFC3,
][]{Kimble2008} in the frame of the HST Proposal GO-13418 (PI:
D. Apai). We acquired the observations in filters F125W
($\lambda_{\mbox{pivot}}$ = 1245.9 nm, full width at half maximum
(FWHM) = 301.5 nm) and F160W ($\lambda_{\mbox{pivot}}$ 1540.52, FWHM =
287.9 nm), roughly corresponding to the J and H bands. The WFC3 pixel
scale is $\approx$13mas. We used the $256\times256$ pixels sub-array
mode to avoid memory dumps during the observations.  In order to
provide a near-continuous coverage for detecting modulations we
observed the 2M1207 system in six consecutive HST orbits, obtaining
data with cadence of $\sim1.5$ minutes over a baseline of 8 hours and
40 minutes. The observations were interrupted by 58 minutes long Earth
occultations every 94 minutes.

The observations applied space craft rolls between each two orbits to
allow roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbits 1, 3, and 5, and those taken in
orbits 2, 4, and 6 differed by $25^{\circ}$. At the separation of
2M1207b this angle difference corresponds to a displacement of
$0.34''$ or 2.75 and 2.30 resolution elements in F125W and F160W,
respectively. In each orbit we took thirteen SPARS10 non-destructive
read-outs with NSAMP=10, alternating between F160W and F125W filters,
with 2--3 identical exposures in one exposure sequence. To improve
sampling and reduce the risk that point spread funciton (PSF) lays on
bad pixels, we applied a 4-point dither pattern with differential "X/Y"
offsets of 1.375" in the detector frame, providing optimal
non-integral (half pixel) stepping of 10.5 and 8.5 pixels in F125W and
F160W, respectively. Over the 6 orbits, we obtained 70 images with 10
non-destructive read-outs in F125W and 64 images in F160W with
exposure time of 88.4~s in each filters.

@@

%The small angular separation of 2M1207 A and B (as shown in Figure
%\ref{fig:1}) makes precise primary star subtraction and photometry
%very difficult. 
%On the under-sampled WFC3 IR detector (plate scale
  %$\sim 0.13''\mbox{pixel}^{-1}$ \cite{dressel2012wide}), the primary
%and the secondary only separate by $\sim 6$ pixels, which is about 5
%times of the FWHM of the PSF. In addition, under-sampling causes
%significant artifacts when shifting the images for registration no
%matter what interpolation method is used.
 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
    variability verification tests. Individual measurements are
    plotted with gray crosses. Photometric measurements of the same exposure
    sequence are binned, and binned photometry are plotted with points
    or squares. Best fitted sinusoidal curves are plotted with solid
    lines. {\em Upper}: binned measurements taken in dithering
    position 1 and 3 (red points) and that taken in 2 and 4 (blue
    squares) are plotted with different symbols. They demonstrate same
    trend of variation. In upper left panel, green line is sinusoidal
    wave fitted with all parameter set free, and purple line is
    sinusoidal wave fitted with period set the same as that of
    F125W. {\em Middle}: sinusoidal waves fitted without using the
    data taken in Orbit \#1. These curves are almost identical to the
    curves plotted in upper panel. {\em Lower}: photometry measured
    with AFEM-added images and best fitted sinusoidal curves. These
    points and curves are also almost identical to those plotted
    in the upper panel.}
  \label{fig:2}
\end{figure*}

@@

For both of 2 rounds, we used {\em Tiny Tim} to calculate 10$\times$
over-sampled model PSFs based on the filters, the spectra 
\citep{Bonnefoy2014, Patience2010}, the telescope's actual focus, and
the telescope jitter.  We used the new set of Tiny Tim parameters
provided by \cite{Biretta2014} to improve model of the cold mask,
OTA spikes, and the coma. The focus parameters were calculated
using the model listed on the STScI
website\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}. To
register \tinytim{} PSF to observed PSF of 2M1207A, we moved the
over-sampled PSF over a coordinate grid (gird size=0.001 pixel) using
cubic interpolation and minimize the rms difference of the
observed and re-binned \tinytim{} PSFs over a region centered on 2M1207A with a
5-pixel-radius aperture centered on 2M1207b excluded. Then we introduce
another \tinytim{} PSF for 2M1207b and  fit the
position of 2M1207 b and the photometries of 2M1207 A and b simultaneously
by least square optimization. From our data set, we discovered that
the difference of observed PSFs and model PSFs were very stable for
given PSF positions. Therefore, at the end of the first round PSF
fitting, we derived 8 (2 roll angles $\times$ 4 dithering positions)
observationally derived correction maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{obs.} - PSF_{model} )}
\end{equation}
where $\mathrm{PSF_{model}}$ is a combination of two \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combined the correction
term linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and fitted the three components together. We found that by including the correction term,
the reduced $\chi^{2}$ of PSF fitting is decreased from $\sim 10$ to
around unity. Relative photometry is acquired from the scaling
parameters of the model PSFs.

PSF profiles change with exposure positions due to
pixelation, especially for the case that WFC3 IR is significantly
under-sampled. Also, the flat fields may potentially have large scale
structures \citep{dressel2012wide}. We found a correlation of
photometry with PSF positions on detector frame for both 2M1207 A and
b. Correction were made by normalizing each group of
exposures that have the same dithering position and roll angle
individually -- we took the median of the fluxes that were measured
from these exposures as normalization factors and divided them from
every photometric measurement. Because the normalization factor for
each group of exposures is calculated across the whole observation,
this normalization step has negligible impacts on variability
analysis.

\subsection{Uncertainty Analysis: White noise}

First we estimated the photon noises for the photometry of 2M1207b. We
propagated the photon noises of every single pixel that were
calculated from count rates and detector gain, to the PSF fitting
results. The photon noises for photometry in F125W and F160W are
1.33\% and 1.02\%, respectively.

Since the PSFs for the 2M1207 A and b were fitted simultaneously in the
PSF photometry procedure, the uncertainties of photometry and position
for the primary and secondary are coupled. Imperfection of position
measurement of 2M1207A could potentially affect the photometry of
2M1207b. We used a Monte Carlo (MC) method to
evaluate the overall systematic of the PSF fitting. We applied the
PSF photometry to images that were added with random Poisson noises and
repeat the photometry procedure for 1000 times. From the distribution of the
result, the uncertainty for F125W and F160W photometry are 
1.34\% and 1.12\%, respectively. We conclude that the white noise of
our observation is dominated by photon noise.

@@

2M1207b were observed at 8 different spots on the
detector (2 rolls $\times$ 4 dithering positions). Imperfect flat
field correction are potential sources of variation as they may
introduce position-dependent differences in the count rates. The
uncertainty of WFC3 IR pipeline flat field is typically $\sim 1\%$
\citep{dressel2012wide}.

In PSF photometry, however, multiple pixels are fitted simultaneously
and we expect a lower than 1\% uncertainty from the flat field
errors. To verify this, we multiplied every image by an artificial flat
field error mask (AFEM) -- a uniformly distributed Gaussian noise array with
mean of 1 and sigma of 1\% -- and repeated the PSF photometry on the
resulting images.  The analysis of these experiments showed almost
identical light curve to the original, verifying that the flat field
errors do not affect our photometry significantly (Figure
\ref{fig:2}).

@@

 As a second test we repeated the analysis neglecting the first
 orbit. The motivation behind this test was that, due to spacecraft
 thermal settling, the first orbit of HST observations is often
 slightly unstable and is often neglected in high-precision studies
 (\textbf{citation?}). Indeed, in our analysis 2M1207 A is
 significantly fainter in the first orbit (Figure \ref{fig:3}) than in
 the subsequent ones.
 Our analysis based on orbits 2--6 only found essentially identical
 results to our analysis of orbits 1--6, based on which we conclude
 that the first less reliable orbit does not affect our results
 significantly (Figure \ref{fig:2}, middle panel).

@@

@@

\label{Results}
We present the first  high-resolution, high-cadence, and
high-precision photometry of a directly imaged planet or
planetary-mass companion. Our observations reveal a modulation in the
light curve of the $\sim 4 \mathrm{M_{{Jup}}}$ companion 2M1207b, the first detection
of modulations in directly imaged planetary mass objects.  The best fit
periods for F125W and F160W are 10.5 and 9.1 hour correspondingly. The
amplitudes for the normalized light curves are 1.45\% and 0.92\% for
F125W and F160W light curves.

@@

The distributions for the periods demonstrate long tail shaped towards
long period, with core region roughly Gaussian. With 64\% confidence,
we estimated the 1-$\sigma$ range for the periods of F125W and F160W
to be $10.5_{-1.2}^{+1.3}$ and $9.1_{-1.0}^{+1.1}$ h,
respectively. The period of best fitted sine wave of F125W light
curve is 1.5h longer than that of F160W, and 0.3--0.4 h larger
1-$\sigma$ standard deviation. We also jointly fit the two band light
curves together forcing the periods of two sinusoids to be the
same. We derived a modulation period of $10.2^{+0.9}_{-0.8}$ h.

We discover that the variation amplitudes in the two bands are significantly
different. By fitting Gaussians to the MC fit result distributions, we
determined that variation amplitude of F125W is 1.45\% with a
standard deviation of 0.22\%, and that of F160W is 0.92\% with a
standard deviation of 0.20\%. The peaks of the two
histograms separate by more than 2-$\sigma$. The variation amplitude
for F125W is 1.58 times of that of F160W light curve.

@@

\begin{figure*}
  \centering
  \plottwo{rotationDiagram}{JH}
  \caption{comparison of 2M1207's rotation period and color change with brown
    dwarfs, \bpic{} b, and solar system planets. {\em Left}: period vs. mass
  plot for 2M1207b (red square), solar system planets and \bpic{} b
  (blue squares), and brown dwarfs (black circles, gray shade). The mass of brown
  dwarfs are assumed to be 30 \mjup{}. The gray rectangle that has
a  $\pm$15 \mjup range in $x$, and a $\pm \sigma$ of brown dwarf
periods range in $y$, indicates a
region where brown dwarfs most likely to appear in this
diagram. Rotation period monotonically decreases with the increase of
mass. {\em Right}: ratio
of variation amplitude in J and H band vs. spectral type for 2M1207b
and brown dwarfs. The colors of the points represent J$-$H magnitude,
and the sizes of the points are proportional to the J band variation
amplitudes. The gray dashed line is the result of a linear fit to
these points. Tight linear correlation of J- and H-band variation ratio
and spectral type is demonstrated in this diagram.}
x \label{fig:5}
\end{figure*}

% A fundamental result of our study is the direct determination of the
% rotation period of a directly imaged planetary-mass object. We
% convert the rotation period to equatorial velocity  by adopting a radius of 1 -- 1.4
% $R_{\mathrm{Jup}}$ for 2M1207b, and 1 $R_{\mathrm{Jup}}$ for field
% brown dwarfs with well defined rotation period from the study of
% \cite{Metchev2015}, and compare their rotation velocities with solar
% system planets and \bpic{} b in the left
% panel of Figure \ref{fig:5}.  The study by
% \citep[][]{Snellen2014} succeeded in measuring \vsini{} for \bpic{} b
% and demonstrated that it fits a trend defined by Solar System planets
% in which more massive planets have faster rotation rates. They suggested
% that this relation is linked to the accretion processes during planet
% formation.

A fundamental result of our study is the direct determination of the
rotation period of a directly imaged planetary mass object. In the
left panel of Figure~\ref{fig:5} we compare the rotation period of
2M1207b to the solar system planets, \bpic b the only other directly
imaged planet with an estimated period and measured {\em v sin i}, and
field brown dwarfs from the study of \citep[][]{Metchev2015}.  The
study by \citep[][]{Snellen2014} succeeded in measuring {\em v sin i}
for \bpic b and demonstrated that it fits a trend defined by Solar
System planets in which more massive planets have faster rotation
rates. The interesting finding that \bpic b, an exoplanet that formed
in a protoplanetary disk, follows this trend suggests a possibly
connection between planet mass, initial angular momentum, and
formation in a disk.
